---
title: "Lab2 - FARS"
author: "Robert Pearson"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, error=TRUE)
library(tidyverse)
```


### Before you proceed:
* Be sure to use SAVE AS to save your own copy of this RMD file. I would remove "starter" from the file name.

* You should already have gone through the R Markdown Demo.rmd file. If you have not, be sure to do so now.

## Additional Setup -- Often this sort of code might not be included or echoed in your document.

### Read FARS data

Read data on all fatal vehicle crashes in Michigan in 2014 from a CSV file (comma-separated-values) into a dataframe. 

```{r further-setup}
FARS_2014_MICH <- read.csv("/cloud/project/FARS_2014_MICH.csv")
```

### Provide your name for captions

**CHANGE** the quoted value below to be *your name*. Keep the object named *myname*. You should also **REPLACE** the value for Author at the very top of this document. 

```{r name-object}
myname <- "Your Name Should Go Here"
```

***

# Categorical Analysis of Fatal Crashes in Michigan in 2014

***

## Part 1 - Complete Code for Graphs Provided


### Bar chart for ROUTE (type of road) using basic ggplot2 syntax

```{r route-basic}
ggplot(FARS_2014_MICH, aes(x=ROUTE)) +
  geom_bar()
```

*I would knit this document now (Knit button above) to create an HTML report. Then look back and forth at the syntax in this markdown file and the generated document, re-knitting as you make changes.*

Notice: This shows counts, and categories are in alphabetic order by default (for a character variable)


### Bar chart for ROUTE showing percents and re-ordering categories

The syntax to show percentages is not very intuitive, so I'm not expecting you to write this from scratch. Later you'll just copy, paste, and modify. I'll explain things for those who are curious. 

1. First, create a vector *routes* with ROUTE values in desired order. This will get used to modify the x axis.
2. `..count..` is a special variable containing the counts for each level of the x variable.
3. `scale_y_continuous` is used here to format values as percentages rather than in the default decimal form (to see its effect you could comment that line out and re-knit). "Continuous" means that the values can fall anywhere on the number line.
4. `scale_x_discrete` uses the vector created above to define possible values of the scale for the x axis and specify their order. "Discrete" means there are a limited number of possible values for this axis.
5. `labs()` adds labels for each axis and the caption

Note you'll get a warning because we didn't provide the value "Unknown" for the x-axis

```{r route-nicer}
routes <- c("Local Street", "County Road", "State Highway", "U.S. Highway", "Interstate")
ggplot(FARS_2014_MICH, aes(x=ROUTE)) +
  geom_bar(aes(y=(..count..)/sum(..count..))) +
  scale_y_continuous(labels=scales::percent) +
  scale_x_discrete(limits=routes) +
  labs(x="Type of Route", y="Relative Frequencies", caption=myname) 
```

#### Insight/Observation/Commentary

**Replace this entire paragraph with a complete sentence or two.** What's something you might tell a family member or friend about this graph if you were to show it to them? Perhaps something striking, surprising, etc. Do not just state a straight fact ("X% of fatal crashes occurred on State Highways")


### Manner of Collision 

Two notes:

* This code should be copied and modified for later plot requests.
* Not Col. Moving = not a collision with a moving vehicle

```{r mancoll}
ggplot(FARS_2014_MICH, aes(x=MAN_COLL)) +
  geom_bar(aes(y=(..count..)/sum(..count..))) +
  scale_y_continuous(labels=scales::percent)+
  labs(x="Manner of Collision", y="Relative Frequencies", caption=myname)
```

*Many of you may not know what to make of the results shown in this graph. I included a separate R script file that you might step through to partially explain things before you write your comments.*

#### Insight/Observation/Commentary

Write a complete sentence or two. 


## Part 2 - Copy/paste/modify code to Produce Graphs

There are three more plots requested below. All of them require the same syntax used for the Manner of Collision plot above. Into each code chunk you should copy and paste that command and then modify the variable name and x-axis label. Be sure to use the exact variable names (including case) as they appear in this names() output. You might delete this paragraph and code block after you've completed everything.

```{r var-names}
names(FARS_2014_MICH)
```



### Did the crash involve Drunk Drivers?

```{r drunk-drivers}

```

#### What's your reaction to this graph?

Write a complete sentence or two.



### Did the crash involve speeding?

```{r speeding}

```

#### What's your reaction to this graph?

Write a complete sentence or two.



### Sex of Driver 

The sex of the driver on form 1 (crashes may involve multiple drivers). This would be unknown for hit-and-runs.

```{r sex}
 
```

#### What's your reaction to this graph?

Write a complete sentence or two.


## Preparing to Submit

When everything is complete, create a PDF document by selecting the arrow next to Knit, and then Knit to PDF. Once the PDF file is produced, you'll need to download it to your computer for uploading to Blackboard. In the Files pane (bottom-right), select the pdf file. Then from the More menu choose Export. 