---
title: "R Markdown Demo"
author: "RPearson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

One of the coolest capabilities of R Studio is the ability to produce reports that contain formatted text, R code, and output produced by R. This is achieved by creating an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. 

When you click the **Knit** button a document will be generated that includes both formatted text and the output of any embedded R code chunks within the document, such as the one below. Click on the **Knit** button now to generate an HTML report. 

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure}
plot(pressure)
```

# Some text formatting examples (This is Header 1)

After knitting the document, look at the syntax in the markdown file and the corresponding formatted output in the generated document.

## Below are types of lists (this is Header 2)

### Unordered lists (this is Header 3)

* item 1
* another un-ordered list item
  + sub-item 1
  + sub-item 2

### Ordered lists (Header 3)

1. An ordered list
1. ordered item 2 with *italic* and **bold** text

## Further reading

For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

***

## Now, open the R Markdown file Lab 2 FARS starter.rmd
